Numerous variables relating to 2016 election in 2021 US GSS^[Code and data are available at: https://github.com/SANGWOONG-LEE/2016electionanalysis.git]

### 2022-03-19 outputs/paper/paper.Rmd
rmd knittable

---
title: "Numerous variables relating to 2016 election in 2021 US GSS"
subtitle: "Possible factors effecting 2016 election"
author: 
  - SangWoong Lee
  - Young Suk 
thanks: "Code and data are available at: https://github.com/SANGWOONG-LEE/2016electionanalysis.git".
date: "`r format(Sys.time(), '%d %B %Y')`"
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
output:
  bookdown::pdf_document2
toc: FALSE
bibliography: references.bib
---

@@

# Data
1.Conveying a sense of the dataset of interest.
2.Dicussion of the survey's methodology and its key features, strengths, weaknesses.
For instance what is the population, frame and sample.
How is the sample recruited 
what sampling approach is taken
what are some of the trade offs of this 
how is non response handled.
(What is good and bad about it)

```{r}
library(ggplot2)
library(dplyr)
reduced_data %>%
  ggplot(aes(x = age, fill = factor(pres16, levels = c("Clinton", "Trump", "Other", "Did not vote")))) +
  geom_bar() +
  scale_fill_manual("Candidate", values = c("Clinton" = "blue", "Trump" = "red", "Other" = "green", "Did not vote" = "yellow")) +
  labs(fil = "Candidate")

reduced_data$pres16 <- factor(reduced_data$pres16, levels = c("Clinton","Trump","Other","Did not vote"))
reduced_data %>%
  ggplot(aes(x = pres16, fill = factor(pres16, levels = c("Clinton", "Trump", "Other", "Did not vote")))) +
  geom_bar() +
  scale_fill_manual("Candidate", values = c("Clinton" = "blue", "Trump" = "red", "Other" = "green", "Did not vote" = "yellow")) +
  labs(fil = "Candidate")

reduced_data %>%
  ggplot(aes(x = gender, fill = gender)) +
  geom_bar() +
  scale_fill_brewer(palette = "Blues")

reduced_data %>%
  ggplot(aes(x = race, fill = race)) +
  geom_bar() +
  scale_fill_brewer(palette = "yellow")

reduced_data %>%
  ggplot(aes(x = gender, fill = factor(pres16, levels = c("Clinton", "Trump", "Other", "Did not vote")))) +
  geom_bar() +
  scale_fill_manual("Candidate", values = c("Clinton" = "blue", "Trump" = "red", "Other" = "green", "Did not vote" = "yellow")) +
  labs(fil = "Candidate")

reduced_data %>%
  ggplot(aes(x = degree, fill = factor(pres16, levels = c("Clinton", "Trump", "Other", "Did not vote")))) +
  geom_bar() +
  scale_fill_manual("Candidate", values = c("Clinton" = "blue", "Trump" = "red", "Other" = "green", "Did not vote" = "yellow")) +
  labs(fil = "Candidate")

reduced_data %>%
  ggplot(aes(x = born_US, fill = factor(pres16, levels = c("Clinton", "Trump", "Other", "Did not vote")))) +
  geom_bar() +
  scale_fill_manual("Candidate", values = c("Clinton" = "blue", "Trump" = "red", "Other" = "green", "Did not vote" = "yellow")) +
  labs(fil = "Candidate")

reduced_data$Family_income <- factor(reduced_data$Family_income, levels = c("Under $10,000","$10,000 ~ $19,999","$20,000 ~ $39,999","$40,000 ~ $59,999","$60,000 ~ $89,999","$90,000 ~ $129,999","$130,000 ~ $169,999","$170,000 or over"))
reduced_data %>%
  ggplot(aes(x = Family_income, fill = factor(pres16, levels = c("Clinton", "Trump", "Other", "Did not vote")))) +
  geom_bar() +
  scale_fill_manual("Candidate", values = c("Clinton" = "blue", "Trump" = "red", "Other" = "green", "Did not vote" = "yellow")) +
  labs(fil = "Candidate") +
  coord_flip()

reduced_data %>%
  ggplot(aes(x = nation_education, fill = factor(pres16, levels = c("Clinton", "Trump", "Other", "Did not vote")))) +
  geom_bar() +
  scale_fill_manual("Candidate", values = c("Clinton" = "blue", "Trump" = "red", "Other" = "green", "Did not vote" = "yellow")) +
  labs(fil = "Candidate")

reduced_data %>%
  ggplot(aes(x = nation_race, fill = factor(pres16, levels = c("Clinton", "Trump", "Other", "Did not vote")))) +
  geom_bar() +
  scale_fill_manual("Candidate", values = c("Clinton" = "blue", "Trump" = "red", "Other" = "green", "Did not vote" = "yellow")) +
  labs(fil = "Candidate")

reduced_data %>%
  ggplot(aes(x = nation_arms, fill = factor(pres16, levels = c("Clinton", "Trump", "Other", "Did not vote")))) +
  geom_bar() +
  scale_fill_manual("Candidate", values = c("Clinton" = "blue", "Trump" = "red", "Other" = "green", "Did not vote" = "yellow")) +
  labs(fil = "Candidate")

reduced_data %>%
  ggplot(aes(x = nation_welfare, fill = factor(pres16, levels = c("Clinton", "Trump", "Other", "Did not vote")))) +
  geom_bar() +
  scale_fill_manual("Candidate", values = c("Clinton" = "blue", "Trump" = "red", "Other" = "green", "Did not vote" = "yellow")) +
  labs(fil = "Candidate")

reduced_data %>%
  ggplot(aes(x = Assistance_to_the_poor, fill = factor(pres16, levels = c("Clinton", "Trump", "Other", "Did not vote")))) +
  geom_bar() +
  scale_fill_manual("Candidate", values = c("Clinton" = "blue", "Trump" = "red", "Other" = "green", "Did not vote" = "yellow")) +
  labs(fil = "Candidate")

reduced_data %>%
  ggplot(aes(x = Assistance_to_other_countries, fill = factor(pres16, levels = c("Clinton", "Trump", "Other", "Did not vote")))) +
  geom_bar() +
  scale_fill_manual("Candidate", values = c("Clinton" = "blue", "Trump" = "red", "Other" = "green", "Did not vote" = "yellow")) +
  labs(fil = "Candidate")

reduced_data %>%
  ggplot(aes(x = social_security, fill = factor(pres16, levels = c("Clinton", "Trump", "Other", "Did not vote")))) +
  geom_bar() +
  scale_fill_manual("Candidate", values = c("Clinton" = "blue", "Trump" = "red", "Other" = "green", "Did not vote" = "yellow")) +
  labs(fil = "Candidate")

reduced_data %>%
  ggplot(aes(x = race, fill = factor(pres16, levels = c("Clinton", "Trump", "Other", "Did not vote")))) +
  geom_bar() +
  scale_fill_manual("Candidate", values = c("Clinton" = "blue", "Trump" = "red", "Other" = "green", "Did not vote" = "yellow")) +
  labs(fil = "Candidate")

reduced_data %>%
  ggplot(aes(x = degree, fill = factor(Family_income, levels = c("$170,000 or over","$130,000 ~ $169,999","$90,000 ~ $129,999","$60,000 ~ $89,999","$40,000 ~ $59,999","$20,000 ~ $39,999","$10,000 ~ $19,999","Under $10,000")))) +
  geom_bar() +
  labs(fill = "Family Income") +
  coord_flip()

```
US demographics: White 76.3%, Black: 13.4%, American Indian: 1.3%, Asian: 5.9%, Hispanic: 18.5%
https://www.census.gov/quickfacts/fact/table/US/PST045221.
Considering the result above, our data sample is fairly recruited.

1. nation_race 
-> This session clearly shows the difference in approval rating between Hillary Clinton and 
Donald Trump. Majority of people who thinks that national spending on improving conditions on
race has to be increased advocated Hillary Clinton. On the other hand, people who did not agree 
with them advocated Donald Trump. I assume that this is due to the image of Donald Trump.

2. nation_arms
-> Interestingly this session shows opposite trend compared to nation_race. Aside from people who
answered "About right" and "NA", the two bars, "Too little" and "Too much" shows us opposite trend.

In general, similar to the indicators above, Hillary Clinton supporters were moved by social equality such as 
"Assistance to the Poor", "nation welfare", "nation education" and "Assistance to the other countries". On the other hand, Donald Trump supporters were interested by more conservative ideas that US is spending too much on the sessions above.

>>>>

---
title: "Numerous variables relating to 2016 election in 2021 US GSS^[Code and data are available at: https://github.com/SANGWOONG-LEE/2016electionanalysis.git]"
subtitle: "Possible factors effecting 2016 election"
author: 
  - SangWoong Lee
  - Young Suk
date: "`r format(Sys.time(), '%d %B %Y')`"
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
output:
  bookdown::pdf_document2
toc: FALSE
bibliography: references.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
library(tidyverse)
library(dplyr)
```

@@

# Data
```{r, echo=FALSE}
reduced_data <- read_csv("outputs/data/prepared_gss.csv")
```
1.Conveying a sense of the dataset of interest.
2.Dicussion of the survey's methodology and its key features, strengths, weaknesses.
For instance what is the population, frame and sample.
How is the sample recruited 
what sampling approach is taken
what are some of the trade offs of this 
how is non response handled.
(What is good and bad about it)

```{r, echo=FALSE}
library(ggplot2)
reduced_data %>%
  ggplot(aes(x = age)) +
  geom_bar(na.rm = T)

reduced_data %>%
  ggplot(aes(x = gender)) +
  geom_bar()

reduced_data %>%
  ggplot(aes(x = race)) +
  geom_bar()

reduced_data %>%
  ggplot(aes(x = gender, fill = pres16)) +
  geom_bar()

reduced_data %>%
  ggplot(aes(x = degree, fill = pres16)) +
  geom_bar()

reduced_data %>%
  ggplot(aes(x = born_US, fill = pres16)) +
  geom_bar()

reduced_data %>%
  ggplot(aes(x = Family_income, fill = pres16)) +
  geom_bar()  + coord_flip()

reduced_data %>%
  ggplot(aes(x = nation_education, fill = pres16)) +
  geom_bar()

reduced_data %>%
  ggplot(aes(x = nation_race, fill = pres16)) +
  geom_bar()

reduced_data %>%
  ggplot(aes(x = nation_arms, fill = pres16)) +
  geom_bar()

reduced_data %>%
  ggplot(aes(x = nation_welfare, fill = pres16)) +
  geom_bar()

reduced_data %>%
  ggplot(aes(x = Assistance_to_the_poor, fill = pres16)) +
  geom_bar()

reduced_data %>%
  ggplot(aes(x = Assistance_to_other_countries, fill = pres16)) +
  geom_bar()

reduced_data %>%
  ggplot(aes(x = social_security, fill = pres16)) +
  geom_bar()

reduced_data %>%
  ggplot(aes(x = race, fill = pres16)) +
  geom_bar()
```
